Software Citation Principles

### 2016-03-01 sc-principles/software-citation-principles.tex
Add 2 notes
Add a note on citation styles

\newcommand{\katznote}[1]{ {\textcolor{blue} { ***DSK: #1 }}}
\newcommand{\niemnote}[1]{ {\textcolor{orange} { ***KEN: #1 }}}
\newcommand{\asnote}[1]{ {\textcolor{red} { ***AMS: #1 }}}
\newcommand{\flnote}[1]{ {\textcolor{olive} { ***FL: #1 }}}

@@

In \cite{ssi-citation-files}, Wilson suggests that software authors include a `CITATION' file that documents exactly how the
authors of the software would like to be cited by others. While this isn't a formal metadata specification (it's not machine
readable for example) this does offer a solution for authors wishing to give explicit instructions to potential citing authors.
In a later post on the SSI blog, Jackson \cite{ssi-how-shalt-i-cite-thee} gives a good overview of some of the approaches
package authors have taken to automate the generation of citation entities such as bibtex entries.

@@

\asnote{Should also discuss the WSSSPE \cite{wssspe} series of workshops in this section}

@@

\begin{enumerate}
\item \textbf{Importance}:
Software should be considered a legitimate, citable product of research. Software citations should be accorded the same importance in the scholarly record as citations of other research objects, such as publications and data.
\item \textbf{Credit and Attribution}
\begin{enumerate}
\item Software citations should facilitate giving scholarly credit and normative and legal attribution to all contributors to the software, recognizing that a single style or mechanism of attribution may not be applicable to all software.
\item In scholarly literature and other products, whenever and wherever a claim relies upon software, the corresponding software should be cited.
\end{enumerate}
\item \textbf{Unique Identification}
A software citation should include a persistent method for identification that is machine actionable, globally unique, and widely used by a community.
\item \textbf{Access}
Software citations should facilitate access to the software themselves and to such associated metadata, documentation, data, and other materials, as are necessary for both humans and machines to make informed use of the referenced software.
\item \textbf{Persistence}
Unique identifiers, and metadata describing the software, and its disposition, should persist -- even beyond the lifespan of the software they describe.
\item \textbf{Specificity and Verifiability}
Software citations should facilitate identification of, access to, and verification of the specific software that support a claim. Citations or citation metadata should include information about provenance and fixity sufficient to facilitate verifying that the specific version and/or portion of software retrieved subsequently is the same as was originally cited.
\item \textbf{Interoperability and Flexibility}
Software citation methods should be sufficiently flexible to accommodate the variant practices among communities, but should not differ so much that they compromise interoperability of software citation practices across communities.
\end{enumerate}

>>>>

\definecolor{darkgreen}{rgb}{0,0.6,0}
\newcommand{\katznote}[1]{ {\textcolor{blue} { ***DSK: #1 }}}
\newcommand{\niemnote}[1]{ {\textcolor{orange} { ***KEN: #1 }}}
\newcommand{\asnote}[1]{ {\textcolor{red} { ***AMS: #1 }}}
\newcommand{\flnote}[1]{ {\textcolor{olive} { ***FL: #1 }}}
\newcommand{\scnote}[1]{ {\textcolor{darkgreen} { ***SCC: #1 }}} % Sou-Cheng Choi

@@

In \cite{ssi-citation-files}, Wilson suggests that software authors include a `CITATION' file that documents exactly how the
authors of the software would like to be cited by others. While this isn't a formal metadata specification (it's not machine
readable for example) this does offer a solution for authors wishing to give explicit instructions to potential citing authors.
In a later post on the SSI blog, Jackson \cite{ssi-how-shalt-i-cite-thee} gives a good overview of some of the approaches
package authors have taken to automate the generation of citation entities such as bibtex entries.
\scnote{So does \cite{knepley2013accurately}}

@@

\asnote{Should also discuss the WSSSPE \cite{wssspe} series of workshops in this section}
\scnote{Section 5.8 and Appendix J in \cite{WSSSPE3} has some preliminary work and discussion on the topic.}

@@

\begin{enumerate}
\item \textbf{Importance}:
Software should be considered a legitimate, citable product of research. Software citations should be accorded the same importance in the scholarly record as citations of other research objects, such as publications and data.
\item \textbf{Credit and Attribution}
\begin{enumerate}
\item Software citations should facilitate giving scholarly credit and normative and legal attribution to all contributors to the software, recognizing that a single style or mechanism of attribution may not be applicable to all software.\scnote{At least four styles (AMS, APA, Chicago, MLA) for software citation are illustrated with examples in ~\cite{lipson2011cite}.}
\item In scholarly literature and other products, whenever and wherever a claim relies upon software, the corresponding software should be cited.
\end{enumerate}
\item \textbf{Unique Identification}
A software citation should include a persistent method for identification that is machine actionable, globally unique, and widely used by a community.
\item \textbf{Access}
Software citations should facilitate access to the software themselves and to such associated metadata, documentation, data, and other materials, as are necessary for both humans and machines to make informed use of the referenced software.
\item \textbf{Persistence}
Unique identifiers, and metadata describing the software, and its disposition, should persist -- even beyond the lifespan of the software they describe.
\item \textbf{Specificity and Verifiability}
Software citations should facilitate identification of, access to, and verification of the specific software that support a claim. Citations or citation metadata should include information about provenance and fixity sufficient to facilitate verifying that the specific version and/or portion of software retrieved subsequently is the same as was originally cited.
\item \textbf{Interoperability and Flexibility}
Software citation methods should be sufficiently flexible to accommodate the variant practices among communities, but should not differ so much that they compromise interoperability of software citation practices across communities.
\end{enumerate}
